---
title: "Love Actually: A Definitive Network Analysis"
format: 
  revealjs:
    theme: simple
    smaller: true
    scrollable: true
    slide-number: true
    code-fold: true
    code-overflow: wrap
    fig-align: center
    fig-width: 9
    fig-height: 5
    embed-resources: true
---

## Introduction

- *Love Actually* (2003) is an ensemble romantic comedy set in London during the month leading up to Christmas.

- The film weaves together nine loosely interconnected love stories, each examining a different facet of love -- the one emotion common to all people.

- Because the film is structured as a set of parallel narratives connected by points of intersection, its character relationships are well represented as a network, motivating my network analysis approach.

---

## Questions of Interest

My project focuses on two related questions:

- **Scene appearances:** When does each character appear across the 71 scenes of the film, and what does the rhythm of their screen time look like?

- **Character networks:** Which characters share scenes with which others, and do those co-appearances cluster into the distinct storylines the film is built around?

---

## Data

I downloaded two CSV files from GitHub and then tidied the data in R for my analysis:

- [**`appearances`**](https://github.com/fivethirtyeight/data/blob/master/love-actually/love_actually_appearances.csv): a table indicating which characters appears in each of the scenes, plus a short description of each scene.

- [**`adjacencies`**](https://github.com/fivethirtyeight/data/blob/master/love-actually/love_actually_adjacencies.csv): a 14 × 14 co-appearance matrix. Diagonal entries store each character's total scene count; off-diagonal entries form the edge weights of the character network.

From the adjacency matrix I derived two data frames: **nodes** (one per character, weighted by total scenes) and **edges** (one per character pair who share a scene, weighted by count).

---

## Tidying the appearances data

```{r}
library(readr)
library(tidyverse)

appearances <- read_csv("data/love_actually_appearances.csv")

appearances <- appearances |>
  ## only keep rows with data, after 71 everything is NA b/c there are no more scenes
  slice(1:71) |>
  mutate(scene_number = row_number()) |>
  pivot_longer(cols = 2:15, names_to = "actors", values_to = "appearances") |>
  mutate(appearances = if_else(appearances == "TRUE", 1, 0)) |>
  replace_na(list(appearances = 0)) |>
  mutate(appearances = as.factor(appearances),
         actors = gsub("_", " ", actors),
         actors = str_to_title(actors))

appearances
```

I pivoted the table to create a long-format table with one row per (character, scene) pair and a binary indicator (1 if character appears, 0 otherwise).

---

## Scene appearances across the film

```{r}
#| code-fold: true
scenes <- appearances |>
  filter(appearances == 1) |>
  arrange(actors, scene_number) |>
  group_by(actors) |>
  mutate(next_scene = lead(scene_number)) |>
  ungroup() |>
  filter(!is.na(next_scene))

order_of_appearance <- appearances |>
  filter(appearances == 1) |>
  group_by(actors) |>
  summarize(first_scene = min(scene_number)) |>
  arrange(desc(first_scene)) |>
  pull(actors)

scenes <- scenes |>
  mutate(actors = factor(actors, levels = order_of_appearance))

appearance_points <- appearances |>
  filter(appearances == 1) |>
  mutate(actors = factor(actors, levels = order_of_appearance))
```

```{r}
#| fig-cap: "Each dot is a scene the character appears in; segments connect first appearance to alst. Characters are ordered by first appearance (top = earliest)."
library(plotly)

s <- ggplot() +
  geom_segment(data = scenes,
               aes(x = scene_number, xend = next_scene,
                   y = actors, yend = actors), color = "magenta1") +
  geom_point(data = appearance_points, 
             aes(x = scene_number, y = actors, 
                 text = paste(scene_number)), 
             color = "red1") +
  labs(title = "All Scene Appearances",
       x = "Scene Number",
       y = NULL) +
  theme_minimal()

ggplotly(s, tooltip = "text")
```

---

## Tidying the adjacencies data

```{r}
adjacencies <- read_csv("data/love_actually_adjacencies.csv")

actors_long <- adjacencies |>
  pivot_longer(cols = -actors,
               names_to = "actor2",
               values_to = "weight") |>
  rename(actor1 = actors) |>
  filter(!is.na(weight)) |>
  mutate(actor1 = str_to_title(gsub("_", " ", actor1)),
         actor2 = str_to_title(gsub("_", " ", actor2)))

adjacencies
```

I pivoted the matrix into a long-format edge table — one row per character pair, with `weight` as the number of shared scenes.

---

## Building nodes and edges

```{r}
library(tidygraph)
library(ggraph)

nodes <- actors_long |>
  filter(actor1 == actor2) |>
  dplyr::select(name = actor1, total_scenes = weight) |>
  mutate(community = case_when(
    name %in% c("Hugh Grant", "Emma Thompson", "Liam Neeson",
                "Alan Rickman", "Heike Makatsch", "Rowan Atkinson") ~ "community 1",
    name %in% c("Kris Marshall", "Abdul Salis", "Colin Firth",
                "Keira Knightley") ~ "community 2",
    .default = "floaters"
  ))

edges <- actors_long |>
  filter(actor1 != actor2, weight > 0) |>
  dplyr::select(from = actor1, to = actor2, weight)
```

:::: {.columns}

::: {.column width="50%"}
**Nodes**
```{r}
nodes
```
:::

::: {.column width="50%"}
**Edges**
```{r}
edges
```
:::

::::

- **Nodes** come from the diagonal: each character + total scene count.
- **Edges** come from the off-diagonal: each shared-scene pair, weighted.

---

## Character network

```{r}
#| fig-cap: "Nodes sized by total scenes; edges weighted by shared-scene count. Communities assigned manually based on storyline membership (to be refined with community-detection algorithms)."
#| fig-width: 10
#| fig-height: 6
network_obj <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)

ggraph(network_obj, layout = "fr") +
  geom_edge_link(aes(width = weight), alpha = 0.4, color = "snow4", show.legend = FALSE) +
  geom_node_point(aes(color = community, size = total_scenes), show.legend = FALSE) +
  geom_node_text(aes(label = gsub(" ", "\n", name)), size = 2.5) +
  scale_edge_width(range = c(0.5, 3)) +
  scale_size(range = c(15, 22)) +
  scale_color_manual(values = c("community 1" = "turquoise1", 
                                "community 2" = "yellow", 
                                "floaters" = "chartreuse")) +
  theme_void() 
```

---

