---
title: "Initial Visualizations"
format: html
---

Users can select scene 1 up to scene x.
Summary of scenes/timeline maybe for what's happened in the movie up to scene x?
For each group of scenes, network graph of characters (relative to movie happenings up to scene x).
Can click on node of network graph to get character description, or branch (connection) to get description of relationships.
Network graph will have branches based on romantic, familial, coworker/acquaintance, friend relationships.

Tidy appearances data.
```{r}
library(readr)
library(tidyverse)

appearances <- read_csv("data/love_actually_appearances.csv")

appearances <- appearances |>
  ## only keep rows with data, after 71 everything is NA b/c there are no more scenes
  slice(1:71) |>
  mutate(scene_number = row_number()) |>
  pivot_longer(cols = 2:15, names_to = "actors", values_to = "appearances") |>
  mutate(appearances = if_else(appearances == "TRUE", 1, 0)) |>
  replace_na(list(appearances = 0)) |>
  mutate(appearances = as.factor(appearances),
         actors = gsub("_", " ", actors),
         actors = str_to_title(actors))
```

Make segment/double-lollipop plot of scene appearances throughout the movie.
```{r}
scenes <- appearances |>
  filter(appearances == 1) |>
  arrange(actors, scene_number) |>
  group_by(actors) |>
  mutate(next_scene = lead(scene_number)) |> # next scene they appear in
  ungroup() |>
  filter(!is.na(next_scene))    

order_of_appearance <- appearances |>
  filter(appearances == 1) |>
  group_by(actors) |>
  summarize(first_scene = min(scene_number)) |>
  arrange(desc(first_scene)) |> 
  pull(actors)

scene_appearances <- scene_appearances |>
  mutate(actors = factor(actors, levels = order_of_appearance))

appearance_points <- appearances |>
  filter(appearances == 1) |>
  mutate(actors = factor(actors, levels = order_of_appearance))

ggplot() +
  geom_segment(data = scenes,
               aes(x = scene_number, xend = next_scene,
                   y = actors, yend = actors), color = "olivedrab") +
  geom_point(data = appearance_points, 
             aes(x = scene_number, y = actors), color = "firebrick") +
  labs(title = "All Scene Appearances",
       x = "Scene Number",
       y = NULL) +
  theme_minimal()
```


Column plot of actors' total number of appearances (scenes appeared in) out of the 71 scenes in Love Actually.
```{r}
most_scenes_order <- appearances |>
  mutate(actors = fct_reorder(actors, appearances == "1", sum)) |>
  group_by(actors) |>
  filter(appearances == "1") |>
  summarise(num_scenes = n())
  

ggplot(most_scenes_order, aes(x = actors, y = num_scenes)) +
  geom_col(fill = "seashell2", color = "lavenderblush3") +
  coord_flip() +
  geom_text(aes(label = num_scenes), hjust = -0.5) +
  theme_minimal() +
  labs(title = "Total Appearances",
       x = NULL,
       y = "Number of Scenes")
```


Tidy adjacencies data.
```{r}
adjacencies <- read_csv("data/love_actually_adjacencies.csv")

actors_long <- adjacencies |>
  pivot_longer(cols = -actors,
               names_to = "actor2",
               values_to = "weight") |>
  rename(actor1 = actors) |>
  filter(!is.na(weight)) |>
  mutate(actor1 = str_to_title(gsub("_", " ", actor1)),
         actor2 = str_to_title(gsub("_", " ", actor2)))
```


Build the nodes and edges data.
```{r} 
library(tidygraph)
library(ggraph)

nodes <- actors_long |>
  filter(actor1 == actor2) |>
  dplyr::select(name = actor1, total_scenes = weight) |>
  mutate(community = case_when(
    name %in% c("Hugh Grant", "Emma Thompson", "Liam Neeson", 
                "Alan Rickman", "Heike Makatsch", "Rowan Atkinson") ~ "community 1",
    name %in% c("Kris Marshall", "Abdul Salis", "Colin Firth", "Keira Knightley") ~ "community 2",
    .default = "floaters"
  ))

edges <- actors_long |>
  filter(actor1 != actor2, weight > 0) |>
  dplyr::select(from = actor1, to = actor2, weight)
```


Build network object and create network graph.
```{r}
network_obj <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)

ggraph(network_obj, layout = "fr") +
  geom_edge_link(aes(width = weight), alpha = 0.4, color = "rosybrown4", show.legend = FALSE) +
  geom_node_point(aes(color = community, size = total_scenes), show.legend = FALSE) +
  geom_node_text(aes(label = gsub(" ", "\n", name)), size = 2.5) +
  scale_edge_width(range = c(0.5, 3)) +
  scale_size(range = c(15, 22)) +
  scale_color_manual(values = c("community 1" = "mistyrose", "community 2" = "plum3", "floaters" = "honeydew")) +
  theme_void() 
```








